---
title: "Crossanova on pharmacokinetic parameter"
author: "Sopan Raman Patil"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction
Here the data corresponds to calculation of pharmacokinetic for dataset
we read the dataset

```{r}
rm(list = ls(all=TRUE))
data=read.csv("D:/SOPAN/DATASET/BABE.csv",header = F)
head(data)
```

- $c_{max}$: maximum concentration
- $t_{max}$: time point at which concentration is maximum
- $AUC_{0-t}$: partial area under curve truncated at the population median of   t max 
- ke rate constant
- $t_{1/2}$: elimination half time 
- $AUC_{0-\infty}$: Area under curve zero to infinity

```{r}
ti=data[1,5:20]
ci=data[2:49,5:20]
pkpara=function(ti,ci){
  pkpara=data.frame("tmax"=0,"cmax"=0,"auct"=0,"aucinf"=0,"thalf"=0,"ke"=0)
  n=length(ti)
  pkpara$cmax=max(ci)     
  a=which.max(ci)
  pkpara$tmax=ti[a]      
  pkpara$auct=sum((ci[2:n]+ci[1:(n-1)])*(ti[2:n]-ti[1:(n-1)]))
  pkpara$ke=-2.303*coefficients(lm(log10(ci[(a+1):n])~ti[(a+1):n]))[2] 
  pkpara$thalf=0.693/pkpara$ke   
  pkpara$aucinf=pkpara$auct+ci[n]/pkpara$ke
  return(pkpara)
}
out=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[1,])))
for(i in 2:48){
  out[i,]=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[i,])))
}
head(out)
```


```{r}
data_1=read.csv("D:/SOPAN/DATASET/BABE2.csv")
df=data.frame(data_1,out)
head(df)
```

```{r}
crosanova=function(Y,Seq,Sub,Per){
  df=data.frame(Y,Seq,Sub,Per)
  Y=df[,1]
  n=length(Y)
  sst=((n-1)/n)*var(Y)
  y11=df[df$Per==1&df$Seq==1,1];n1=length(y11)
  y12=df[df$Per==1&df$Seq==2,1]
  y21=df[df$Per==2&df$Seq==1,1]
  y22=df[df$Per==2&df$Seq==2,1];n2=length(y22)
  df1=data.frame(y11,y21)
  df2=data.frame(y12,y22)
  v1=apply(df1,1,var)
  v2=apply(df2,1,var)
  sswit=sum(v1)+sum(v2)
  m1=sum((apply(df1,1,mean)-mean(Y))^2)
  m2=sum((apply(df2,1,mean)-mean(Y))^2)
  ssbet=2*(m1+m2)
  u1=y11+y21;u2=y12+y22
  d1=(y21-y11)/2;d2=(y22-y12)/2
  o1=d1;o2=d2
  
  chat=mean(u2)-mean(u1)
  fhat=mean(d1)-mean(d2)
  phat=mean(o1)-mean(o2)
  sscarry=2*n1*n2/(n1+n2)*chat^2/4
  ssdrug=2*n1*n2/(n1+n2)*fhat^2
  ssperiod=2*n1*n2/(n1+n2)*phat^2
  ssintra=sswit-ssdrug-ssperiod
  ssinter=ssbet-sscarry
  
  source1=c("carry","inter","drug","period","intra","Total") 
  dof=c(1,n1+n2-2,1,1,n1+n2-2,n-1)
  ss=c(sscarry,ssinter,ssdrug,ssperiod,ssintra,sst)
  msq=ss/dof
  fcal=rep(0,6)
  fcal[1]=msq[1]/msq[2]
  fcal[2:4]=msq[2:4]/msq[5]
  pval=rep(0,6)
  pval[1]=1-pf(fcal[1],dof[1],dof[2])
  pval[2:4]=1-pf(fcal[2:4],dof[2:4],dof[5])
  return(data.frame("s_var"=source1,"DF"=dof,"sof_squ"=round(ss,4),"Fcal"=round(fcal,4),"p_value"=round(pval,4)))
}

```

# Anova corresponds to concentration

```{r}

crosanova(df$cmax,df$Seq,df$Sub,df$Per)

```

# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_{04}$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present.

# Anova corresponds to partial auc truncated at the population median of t max

```{r}
crosanova(df$auct,df$Seq,df$Sub,df$Per)
```

# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to  0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here   from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0     here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present.

# Anova corresponds to rate constant

```{r}
crosanova(df$ke,df$Seq,df$Sub,df$Per)
```

# Conclusion

- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to   0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here   from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0     here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present.

# Anova corresponds to area under curve zero to infinity 

```{r}
crosanova(df$aucinf,df$Seq,df$Sub,df$Per)
```

# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs H$H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present.

# Anova corresponds to elimination half time 

```{r}
crosanova(df$thalf,df$Seq,df$Sub,df$Per)
```

# Conclusion

- carryover effect $H_01:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present.
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present.

# plot of time point and concetration

```{r}
time=unlist(ti)
y1=apply(data[data$V3==1,5:20],2,mean)
x1=plot(time,y1,type="l",col='blue',xlab = "time",ylab="concentration")
par(new=TRUE)
y2=apply(data[data$V3==2,5:20],2,mean)
x2=plot(time,y2,type="l",col='red',xlab = "time",ylab="concentration")
title(main = "Area under curve", xlab = "time", ylab = "concentration")

```




#from graph we have seen that there is Bioavailability and bioequivalence 
