---
title: "PKpara and Effect estimation"
author: "Patil Sopan Raman"
date: "2022-11-07"
output:
  word_document: default
  html_document:
    df_print: paged
---

# Introduction
Here the data corresponds to calculation of pharmacokinetic for dataset
we read the dataset

# Data

```{r}
rm(list = ls(all=TRUE))
data=read.csv("D:/SOPAN/DATASET/pkparadata.csv",header = F)
```

# Seperating time point and concetration corresponding to each subject

```{r}
ti=data[1,5:20]
ci=data[2:49,5:20]
```
# Define function pkpara finding pharmacokinetics 

 
- $c_{max}$: maximum concentration
- $t_{max}$: time point at which concentration is maximum
- $AUC_{0-t}$: partial area under curve truncated at the population median of t max 
- ke rate constant
- $t_{1}/{2}$: elimination half time 
- $AUC_{0-\infty}$: Area under curve zero to infinity



```{r}
pkpara=function(ti,ci){
  pkpara=data.frame("tmax"=0,"cmax"=0,"auct"=0,"aucinf"=0,"thalf"=0,"ke"=0)
  n=length(ti)
  pkpara$cmax=max(ci)     
  a=which.max(ci)
  pkpara$tmax=ti[a]      
  pkpara$auct=sum((ci[2:n]+ci[1:(n-1)])*(ti[2:n]-ti[1:(n-1)]))
  pkpara$ke=-2.303*coefficients(lm(log10(ci[(a+1):n])~ti[(a+1):n]))[2] 
  pkpara$thalf=0.693/pkpara$ke   
  pkpara$aucinf=pkpara$auct+ci[n]/pkpara$ke
  return(pkpara)
}
```

# pharmacokinetic for whole data set

```{r}
out=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[1,])))
for(i in 2:48){
  out[i,]=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[i,])))
}
out
```
# Five point summary of pharmacokinetic parameter

```{r}
summary(out)
```
# Data of standard 2x2 crossover design

```{r}

data_1=read.csv("D:/SOPAN/DATASET/BABEData2.csv")

```
# Statistical inferences for the fixed effects

- here we calculate carryover, drug formulation effect, period effect
- $C_{hat}$: carryover effect (estimated)
- $F_{hat}$: linear contrast of the seq by period means
- $P_{hat}$: unbiased estimator of period effect

```{r}
y11=data_1[data_1$Per==1&data_1$Seq==1,1];n1=length(y11)
y12=data_1[data_1$Per==1&data_1$Seq==2,1]
y21=data_1[data_1$Per==2&data_1$Seq==1,1]
y22=data_1[data_1$Per==2&data_1$Seq==2,1];n2=length(y22)
U1=y11+y21   # subject total w.r.t. sequence 1
U2=y12+y22    # subject total w.r.t. sequence 2
D1=(y21-y11)/2   # period differences for each subject within each sequence 1
D2=(y22-y12)/2   # period differences for each subject within each sequence 2
O1=D1
O2=D2
sd2u=((n1-1)*var(U1)+(n2-1)*var(U2))/(n1+n2-2)  # pooled estimate of U
sd2d=((n1-1)*var(D1)+(n2-1)*var(D2))/(n1+n2-2)  # pooled estimate of D
chat=mean(U2)-mean(U1)  
fhat=mean(D1)-mean(D2) 
phat=mean(O1)-mean(O2)   
vchat=sd2u*(1/n1+1/n2)   # pooled sample variance of carryover effect
vfhat=sd2d*(1/n1+1/n2)   # pooled sample variance of drug formulation effect
effect=c("carry","drug","perid")
mvue=c(chat,fhat,phat)
vhat=c(vchat,vfhat,vfhat)
test_stat=mvue/sqrt(vhat)
pv=2*(1-pt(abs(test_stat),n1+n2-2))
df=data.frame(effect,round(mvue,2),ss=round(vhat,2),f_ratio=round(test_stat,3),pv=round(pv,4))
df
```
# Conclusion

- carryover effect: $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:C_T-C_R$ not equal to 0 here from p_value we fail to reject $H_{01}$  at 5% level of significance.
- Drug formulation effect: $H_{02}:{F_T}-{F_R}=0$ vs $H_{12}:F1-F2$ not equal to 0 here from p_value we fail to reject $H_{02}$  at 5% level of significance.
- Period effect: $H_{03}:{P_1}-{P_2}=0$ vs $H_{13}:{P_1}-{P_2}$ not equal to 0, here from p_value we fail to reject $H_{03}$ at 5% level of significance.

# Data of standard 2x2 crossover design

```{r}
data_1=read.csv("D:/SOPAN/DATASET/BABEData2.csv")

```
# Anova for standard 2x2 crossover design

```{r}
n=length(data_1$Y)
y11=data_1[data_1$Per==1&data_1$Seq==1,1];n1=length(y11)
y12=data_1[data_1$Per==1&data_1$Seq==2,1]
y21=data_1[data_1$Per==2&data_1$Seq==1,1]
y22=data_1[data_1$Per==2&data_1$Seq==2,1];n2=length(y22)
ss_total=((n-1)/n)*var(data_1[,1])
df=data.frame(c(y11,y12),c(y21,y22))
ss_within=sum(apply(df,1,var))
n1=length(y11)
n2=length(y22)
ss_between=2*sum((apply(df,1,mean)-mean(data_1$Y))^2)
sstotal=(n-1)*var(data_1$Y)
chat=(mean(y12)+mean(y22))-(mean(y11)+mean(y21))
sscarry=2*n1*n2/(n1+n2)*chat^2/4
ssinter=ss_between-sscarry
fhat=(((mean(y21)-mean(y11))-(mean(y22)-mean(y12)))/2)^2
ssdrug=2*n1*n2/(n1+n2)*fhat
phat=(((mean(y21)-mean(y11))-(mean(y12)-mean(y22)))/2)^2
ssperiod=2*n1*n2/(n1+n2)*phat

ssintra=ss_within-ssdrug-ssperiod
source1=c("carry","inter","drug","period","intra","total")
df=c(1,(n1+n2-2),1,1,(n1+n2-2),2*(n1+n2)-1)
ss=c(sscarry,ssinter,ssdrug,ssperiod,ssintra,sstotal)
msq=ss/df
fcal=rep(0,6)
fcal[1]=msq[1]/msq[2]
fcal[2:4]=msq[2:4]/msq[5]
pv=1-pt(msq,1,n1+n2-2)
data.frame("source"=source1,"df"=df,"sumofsq"=ss,mss=round(msq,3),"fcal"=round(fcal,4),"p_val"=pv)

```
# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{01}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is not present.
- Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{14}:{F_1}-{F_2}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{13}:{P_1}-{P_2}$not equal to 0 here from p_value reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is present. 

# Here we combine dataframe of BABEData2 and six pk parameter dataframe

```{r}
df=data.frame(data_1,out)
```

```{r}
crosanova=function(Y,Seq,Sub,Per){
  df=data.frame(Y,Seq,Sub,Per)
  Y=df[,1]
  n=length(Y)
  sst=((n-1)/n)*var(Y)
  y11=df[df$Per==1&df$Seq==1,1];n1=length(y11)
  y12=df[df$Per==1&df$Seq==2,1]
  y21=df[df$Per==2&df$Seq==1,1]
  y22=df[df$Per==2&df$Seq==2,1];n2=length(y22)
  df1=data.frame(y11,y21)
  df2=data.frame(y12,y22)
  v1=apply(df1,1,var)
  v2=apply(df2,1,var)
  sswit=sum(v1)+sum(v2)
  m1=sum((apply(df1,1,mean)-mean(Y))^2)
  m2=sum((apply(df2,1,mean)-mean(Y))^2)
  ssbet=2*(m1+m2)
  u1=y11+y21;u2=y12+y22
  d1=(y21-y11)/2;d2=(y22-y12)/2
  o1=d1;o2=d2
  
  chat=mean(u2)-mean(u1)
  fhat=mean(d1)-mean(d2)
  phat=mean(o1)-mean(o2)
  sscarry=2*n1*n2/(n1+n2)*chat^2/4
  ssdrug=2*n1*n2/(n1+n2)*fhat^2
  ssperiod=2*n1*n2/(n1+n2)*phat^2
  ssintra=sswit-ssdrug-ssperiod
  ssinter=ssbet-sscarry
  
  source1=c("carry","inter","drug","period","intra","Total") 
  dof=c(1,n1+n2-2,1,1,n1+n2-2,n-1)
  ss=c(sscarry,ssinter,ssdrug,ssperiod,ssintra,sst)
  msq=ss/dof
  fcal=rep(0,6)
  fcal[1]=msq[1]/msq[2]
  fcal[2:4]=msq[2:4]/msq[5]
  pval=rep(0,6)
  pval[1]=1-pf(fcal[1],dof[1],dof[2])
  pval[2:4]=1-pf(fcal[2:4],dof[2:4],dof[5])
  return(data.frame("s_var"=source1,"DF"=dof,"sof_squ"=round(ss,4),"Fcal"=round(fcal,4),"p_value"=round(pval,4)))
}

```
# Anova corresponds to concentration
```{r}
crosanova(df$cmax,df$Seq,df$Sub,df$Per)
```
# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_{04}$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present. 

# Anova corresponds to time point

```{r}
crosanova(df$tmax,df$Seq,df$Sub,df$Per)
```
# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present. 

# Anova corresponds to partial auc truncated at the population median of t max

```{r}
crosanova(df$auct,df$Seq,df$Sub,df$Per)
```
# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present. 

# Anova corresponds to rate constant

```{r}
crosanova(df$ke,df$Seq,df$Sub,df$Per)
```
# Conclusion

- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present. 

# Anova corresponds to area under curve zero to infinity 

```{r}
crosanova(df$aucinf,df$Seq,df$Sub,df$Per)
```
# Conclusion
- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs H$H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present. 
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present.

# Anova corresponds to elimination half time 

```{r}
crosanova(df$thalf,df$Seq,df$Sub,df$Per)
```
# Conclusion

- carryover effect $H_01:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance.
- from p_value we can conclude that inter subject variability is present.
- Drug effect $H_03:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}$ not equal to 0 here from p_value we fail to reject $H_03$  at 5% level of significance.
- period effect $H_04:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}$ not equal to 0 here from p_value fail to reject $H_04$  at 5% level of significance.
- from p_value we can conclude that intra subject variability is not present. 

# plot of time point and concetration

```{r}
time=unlist(ti)
y1=apply(data[data$V3==1,5:20],2,mean)
x1=plot(time,y1,type="l",col='blue',xlab = "time",ylab="concentration")
par(new=TRUE)
y2=apply(data[data$V3==2,5:20],2,mean)
x2=plot(time,y2,type="l",col='red',xlab = "time",ylab="concentration")
title(main = "Area under curve", xlab = "time", ylab = "concentration")

```
#from graph we have seen that there is Bioavailability and bioequivalence 