---
title: "pharmaco parameter"
author: "Sopan Raman Patil"
date: "`r Sys.Date()`"
output: html_document
---

# Pharmacokinetic parameters

- $c_{max}$: maximum concentration
- $t_{max}$: time point at which concentration is maximum
- $AUC_{0-t}$: partial area under curve truncated at the population 
  median of t max 
- ke rate constant
- $t_{1/2}$: elimination half time 
- $AUC_{0-\infty}$: Area under curve zero to infinity

```{r}

time_point=c(0,0.5,1,1.5,2,3,4,6,8,12,16,24,32)
conc=c(0,0,2.8,4.4,4.4,4.7,4.1,4,3.6,3,2.5,2,1.6)
n=length(time_point)
cmax=max(conc)
a=which.max(conc)
tmax=time_point[a]
auc_t=sum((conc[2:n]+conc[1:(n-1)])*(time_point[2:n]-time_point[1:(n-1)]))
model=lm(log10(conc[(a+1):n])~time_point[(a+1):n])
ke=-2.303*coefficients(model)[2]
t_half=0.693/ke
auc_inf=auc_t+conc[n]/ke
plot(time_point,conc,type = 'l',col='black')
```


# Pharmacokinetic parameters for BABE data

```{r}
rm(list = ls(all=TRUE))
data=read.csv("D:/SOPAN/DATASET/BABE.csv",header = F)
head(data)
```

## Here we seperate time point and concentration from the data
```{r}
ti=data[1,5:20]
ci=data[2:49,5:20]
pkpara=function(ti,ci){
  pkpara=data.frame("tmax"=0,"cmax"=0,"auct"=0,"aucinf"=0,"thalf"=0,"ke"=0)
  n=length(ti)
  pkpara$cmax=max(ci)     
  a=which.max(ci)
  pkpara$tmax=ti[a]      
  pkpara$auct=sum((ci[2:n]+ci[1:(n-1)])*(ti[2:n]-ti[1:(n-1)]))
  pkpara$ke=-2.303*coefficients(lm(log10(ci[(a+1):n])~ti[(a+1):n]))[2] 
  pkpara$thalf=0.693/pkpara$ke   
  pkpara$aucinf=pkpara$auct+ci[n]/pkpara$ke
  return(pkpara)
}
```


```{r}
out=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[1,])))
for(i in 2:48){
  out[i,]=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[i,])))
}
head(out)
```
```{r}
summary(out)
```
